The Effects of Stellar Binarity on Transit Survey Occurrence Rates

### 2017-10-03 paper/ms.tex
wordsmithed; did needlessly hard integrals

One might also imagine an apparently-searchable star which in no stellar 
component is of the desired type.
For single stars, this may simply be from erroneous stellar parameters.
In binaries, one might imagine the combined light of two stars somehow making 
them resemble a single star of the desired type.  We will 
not consider errors of this type.  We will assume that all the 
apparently-searchable stars are either single stars of
the desired type, or binaries in which either the primary or else both 
components are of the desired type.

@@

The above approach ignores incompleteness for binary systems.
If we set the limiting flux to give 100\% completeness for single stars, we 
will unwittingly select binaries near that flux limit, for which dilution 
will push transit signals below the detection threshold.
This is because a flux limit maps onto three distinct maximum 
detection distances for a population of single stars and twin binaries:
\begin{enumerate}
\item $d_{{\rm max},s}$: the maximum distance out to which the desired type 
	of planet is detectable around single stars.
	These are the selected ``searchable, single stars''.
%
\item $d_{{\rm max},d}$: the maximum distance out to which double stars
  are selected (down to the same minimum flux as that defining $d_{{\rm 
  max},s}$). 
  This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
  distance out to which planets are detectable around double stars. For a 
  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
  L^{1/2} d^{-1}$,
  \begin{equation}
    d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
    (1 + \gamma_R)^{-1} d_{{\rm max},d}.
  \end{equation}
  For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
  are in fact binary systems can yield detectable planets.

@@

The model of Sec.~\ref{sec:model_1} is of course idealized, but it helps
for developing intuition.
We now let the light ratio $\gamma_R = L_2/L_1$ vary across the population 
of star systems.
It does so because the underlying mass ratio $q=M_2/M_1$ varies.
We keep the primary mass fixed as $M_1$, which is also the mass of all single 
stars.

One parametrization for the distribution of 
binary mass ratios in a volume-limited sample is $f(q)\propto q^\gamma$.
[Duchene and Kraus 2013], fitting all the multiple systems of [Raghavan et al 
(2010)]'s Fig 16, quote $\gamma = 0.28\pm0.05$ for $0.7<M_\star/M_\odot<1.3$.
Examining only the binary systems of Rhagavan et al 2010, Fig 16, the 
distribution is roughly uniform,
\begin{equation}
f(q) =
\Bigg\{\begin{array}{lr}
c_q & 0.1 < q \leq 1  \\
0 & {\rm otherwise},
\end{array}
\label{eq:mass_ratio}
\end{equation}
provided we ignore the peak at high mass ratios for analytic convenience. $c_q 
= 1/9$ is the normalization.
We use this latter mass ratio distribution.

For the mass-luminosity relation, for analytic convenience we assume 
$L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
As a simple check, we fit lines to mass-luminosity data collected by 
Torres et al. [2009] for dwarfs earlier than spectral class M, and Benedict 
et al. [2016] for dwarfs later than spectral class M.
To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
luminosities, we interpolated over E. Mamajek's 
table\footnote{
	\texttt{pas.rochester.edu/\textasciitilde 
		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
	downloaded 2017.08.02}.
In log-log space, we let the intersection point of the lines float, and made 
various cuts on the data as indicated in Fig~\ref{fig:mass_luminosity}. The 
fit parameters are available in a footnote\footnote{
	m lo: 1.8818719873988132~--~
	c lo: -0.9799647314108376~--~
	m hi: 5.1540712426599882~--~
	c hi: 0.0127626185389781~--~
	M at merge: 0.4972991257826812~--~
	L at merge: 0.0281260412126928.
}. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
$L=M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.

The final addition to our model is that we must assume a stellar mass-radius 
relation.
This affects the transit probabilities and signal sizes about the 
secondaries of binary star systems.
[Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
binary data,
\begin{equation}
R = 1.06 M^{0.945}
\label{eq:mass_radius}
\end{equation}
applies for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 
1991's stated bounds), where each quantity is given in solar units. 
Since again, our main purpose is to develop intuition about observational 
biases, we approximate this as $R\propto M$.

@@

Counting the number of desired stars, $N_s$ is the same as in 
Sec.~\ref{sec:model_1}. However
\begin{equation}
N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(\gamma_R)
\end{equation}
for
\begin{equation}
d_{\rm max,d}(\gamma_R) =
d_{\rm max,s}\times (1+\gamma_R)^{1/2},
\end{equation}
where $d_{\rm max,s}$ is given by Eq.~\ref{eq:d_max}.

For a given binary system, $q$, $\gamma_R$, and $d_{\rm 
	max,d}(\gamma_R)$ are all random variables.
Since $N_d$ is a function of $d_{\rm max,d}$, 
the number of double star systems becomes a random variable.
Applying the chain rule for probability density functions, the distribution 
function for the number of binary systems,
${\rm prob}(N_d)$, can be written 
\begin{equation}
{\rm prob}(N_d) = {\rm prob}(q(\gamma_R)) 
\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
\end{equation}

@@

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/joint_prob_r_q.pdf}
	\end{center}
	\caption{The joint probability distribution of a binary star's position and 
		mass ratio in a magnitude-limited sample. This plot assumes a different 
		$M(L)$ relation than previous plots: $L=M^3$. Note that the marginalized 
		mass ratio distribution is uniform at any given distance.}
	\label{fig:joint_prob_r_q}
\end{figure}

@@

The distribution function for the mass ratio of binaries in a magnitude 
limited sample can be found by marginalizing over the joint 
distribution for a binary star's position $r$ and mass ratio $q$.
Explicitly, using `ml' as a subscript for `magnitude-limited',
\begin{equation}
{\rm prob}_{\rm ml}(r,q) = {\rm prob}(q|r){\rm prob}_{\rm ml}(r),
\end{equation} 
where ${\rm prob}(q|r)$ is given by Eq.~\ref{eq:mass_ratio}, and 
\begin{align}
{\rm prob}_{\rm ml}(r) &= \int_{0}^{1} {\rm prob}_{\rm ml}(r|q)\,{\rm 
	prob}(q)\,{\rm d}q
\\
&= \int_{0.1}^{1} \frac{1}{3} \frac{r^2}{d_{\rm max,d}^3(q)}\,{\rm d}q
\quad {\rm if\ }r<d_{\rm max}(q),
\end{align}
and otherwise zero. The resulting joint probability distribution is plotted in 
Fig.~\ref{fig:joint_prob_r_q}.

Marginalizing the joint distribution over distance, and assuming $L\propto 
M^{3.5}$,
\begin{align}
{\rm prob}_{\rm ml}(q) &= 
\int_{0}^{d_{\rm max}(q)} {\rm prob}_{\rm ml}(r,q) \,{\rm d}r \\
&= \mathcal{Z}\cdot (1+q^{7/2})^{3/2}\quad {\rm if\ }0.1<q\leq 1,
\label{eq:pdf_q_mag_limited}
\end{align}
and otherwise 0, for $\mathcal{Z}$ the appropriate normalization constant.
This marginalized distribution~--~the distribution function of mass ratios for 
binaries in a magnitude-limited sample~--~is plotted in 
Fig.~\ref{fig:q_distribn_mag_limited}.

More correctly then, we must write
\begin{equation}
{\rm prob}(N_d) = {\rm prob_{ml}}(q(\gamma_R)) 
\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
\end{equation}

Doing some algebra, and assuming $\gamma_R = q^\alpha$, this can be shown to be
\begin{align}
{\rm prob}(N_d) \propto N_d^{2/3} &\left(\frac{3}{4\pi 
	n_d}\right)^{5/3} d_{\rm max,s}^{-5} \ \nonumber\\
&\times \left[ 
	\left(\frac{3 N_d}{4\pi n_d}\right)^{2/3}  d_{\rm max,s}^{-2} - 1
 \right]^{\frac{1-\alpha}{\alpha} }
\end{align}
over the interval
\begin{equation}
N_d^{\rm lower} = \frac{4 \pi n_d}{3} (\sqrt{0.1^3 + 1} d_{\rm max,s})^3,
\end{equation}
\begin{equation}
N_d^{\rm upper} = \frac{4 \pi n_d}{3} (\sqrt{2}d_{\rm max,s})^3,
\end{equation}
and outside the stated interval ${\rm prob}(N_d) = 0$.

@@

The reason we derived all this is that we wanted to know the number of 
searchable binary stars.
The expected value for the number of double star systems in the sample is
\begin{equation}
\langle N_d \rangle = \int_{0}^{\infty} N_d\, {\rm prob}(N_d) \,{\rm d}N_d.
\end{equation}
Letting $\alpha=3.5$, this can be expressed as
\begin{equation}
\langle N_d \rangle \propto \mathcal{R}^{-3/2}
\int_{0.1^{7/2} + 1}^{2} u^3 (u-1)^{-5/7} \,{\rm d}u ,
\label{eq:integral_Nd}
\end{equation}
where non-dimensional constants are not shown, and $\mathcal{R}$ is defined as
\begin{equation}
\mathcal{R} \equiv \left(\frac{3}{4\pi n_d}\right)^{2/3} d_{\rm max,s}^{-2}.
\end{equation}
The integral in Eq.~\ref{eq:integral_Nd} has a closed-form solution in terms 
of polynomials. Once evaluated, and after 
normalizing ${\rm prob}(N_d)$, to 4 significant digits it gives
\begin{equation}
\langle N_d \rangle = 1.590 \left(\frac{4\pi n_d}{3}\right) d_{\rm max,s}^3.
\end{equation}
This gives the useful expression
\begin{align}
\frac{\langle N_d \rangle}{N_s} &= 1.590\, \frac{n_d}{n_s} \\
&= 1.590 \left(\frac{{\rm BF}}{1-{\rm BF}}\right).
\end{align}
Taking the sun-like star binary fraction of ${\rm BF} = 0.44$, this gives
$\langle N_d \rangle / N_s = 1.25$.
Compared to the binary-twin thought experiment of Sec.~\ref{sec:model_1}, in 
which the SNR-limited sample had $\approx 2.8$ times more binary systems than 
single systems, there are now $\approx 1.25$ times as many.
This is sensible, because the mean light ratio has decreased; it was 
previously 1, it is now $\langle \gamma_R \rangle \approx 0.27$.

@@

Returning to the original question, the total number of stars in the sample is
\begin{equation}
\langle N_{\rm stars,tot} \rangle = N_s + 2 \langle N_d \rangle.
\end{equation}
For stars in binary systems, only a fraction of them are of the desired type.
Defining the ``desired type'' to be $0.7<M_\star/M_\odot<1.3$, this can be 
evaluated by integrating Eq.~\ref{eq:pdf_q_mag_limited} from $0.7M_\odot$ to 
$M_\odot$.
The result is that $47.5\%$ of secondaries are of the desired type.
Thus the expected number of stars of the desired type in the sample is
\begin{equation}
\langle N_{\rm stars} \rangle = N_s + 1.475 \langle N_d \rangle.
\end{equation}

The number of planets of the desired type is then
\begin{equation}
\langle N_{\rm planets} \rangle = 
\Gamma_{t,s}N_s + \Gamma_{t,d}\langle N_d\rangle + 
0.475\Gamma_{t,d}w_{d2} \langle N_d\rangle,
\end{equation}
and the true occurrence rate is
\begin{align}
\Gamma_t &= \frac{\langle N_{\rm planets} \rangle }{\langle N_{\rm stars} 
\rangle} \\
\Gamma_t &= 
\frac{
	\Gamma_{t,s}N_s + (1+0.475w_{d2})\Gamma_{t,d}\langle N_d\rangle
	}{
	N_s + 1.475 \langle N_d \rangle
	}.
\end{align}

@@

The observer who has never heard of binary star systems wants to find the 
occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars 
with mass $0.7<M_\star/M_\odot < 1.3$ (with corresponding stellar radii and 
luminosities).
They detected $N_{\rm det}$ transit signals that appear to be from planets of 
this desired type.
The occurrence rate they then find, {\it i.e.} the number of planets divided 
by number of ``stars'' (really, stellar systems) is
\begin{equation}
\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
\end{equation}

@@

The total number of planet detections of the desired type is 
\begin{equation}
\langle N_{\rm det}\rangle = 
N_{{\rm det},s} 
+ 
\langle N_{{\rm det},d}\rangle.
\end{equation}
As in Sec.~\ref{sec:model_1},
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
\label{eq:N_det_s_model2}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, $f_{s,g}$ is the geometric transit probability, 
and $f_{s,c}$ is the fraction of these transiting planets that are observed 
with signal to noise greater than the minimum detection threshold (the 
completeness). 
For our SNR-limited, varying-$\gamma_R$ survey, $f_{s,g} = R_1/a\equiv f_g$, 
and $f_{s,c}=1$.

For planets detected in binary systems, only a fraction will be ``of the 
desired type'', since only a fraction of the stars are of the desired type.
This is the same factor $f_{\rm desired}$ that we have shown is 0.475. Thus
\begin{equation}
\langle N_{{\rm det},d} \rangle
= 
(1+0.475w_{d2}) \langle N_d\rangle \Gamma_{t,d} f_{d,g} f_{d,c}.
\label{eq:N_det_d_model2}
\end{equation}
Now $(1+0.475w_{d2}) \langle N_d\rangle \Gamma_{t,d}$ is the number of 
planets of the desired type in the double star systems of the sample.
The geometric transit probability $f_{d,g}$ and completeness $f_{d,c}$ require 
more careful evaluation.
First, separate them into their sub-components about primaries ``d1'' and 
secondaries ``d2'':
\begin{align}
\langle N_{{\rm det},d} \rangle
= 
\langle N_d\rangle \Gamma_{t,d} ( f_{d1,g} f_{d1,c} 
+
0.475w_{d2}  f_{d2,g} f_{d2,c}
)
\end{align}

The geometric transit probability $f_{d,g}=R_1/a$ 
for our twin-binary survey, and $f_{d,c}=1/8$ by the geometric argument 
presented in the initial discussion of limiting distances.

\subsection{Errors on derived occurrence rates for Model \#2}

\subsection{How many planets are in the sample?}

@@

>>>>

One might also imagine an apparently-searchable star which in no stellar 
component is of the desired type.
%For single stars, this may simply be from erroneous stellar parameters.
%In binaries, one might imagine the combined light of two stars somehow making 
%them resemble a single star of the desired type.  
We will not consider errors of this type.  We will assume that all the 
apparently-searchable stars are either single stars of
the desired type, or binaries in which either the primary or else both 
components are of the desired type.

@@

The above approach ignores incompleteness for binary systems.
If we set the limiting flux to give 100\% completeness for single stars, we 
will unwittingly select binaries near that flux limit, for which dilution 
will push transit signals below the detection threshold.
This is because a flux limit maps onto three distinct maximum 
detection distances for a population of single stars and twin binaries:
\begin{enumerate}
\item $d_{{\rm max},s}$: the maximum distance out to which the desired type 
	of planet is detectable around single stars.
	These are the selected ``searchable, single stars''.
%
\item $d_{{\rm max},d}$: the maximum distance out to which double stars
  are selected (down to the same minimum flux as that defining $d_{{\rm 
  max},s}$). 
  This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
  distance out to which planets are detectable around double stars. For a 
  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
  L^{1/2} d^{-1}$,
  \begin{equation}
    d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
    (1 + \gamma_R)^{-1} d_{{\rm max},d}.
  \end{equation}
  For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
  are in fact binary systems can yield detectable planets (the ratio of 
  volume in which planets can be detected to the actual selected volume).
  For a more realistic sample, $\gamma_R \approx 0.1$, and there is an 
  interesting asymmetry.
  One finds the completeness for planets which orbit the primary star, 
  $f_{d1,c}$ to be
  \begin{equation}
  f_{d1,c} = (1+\gamma_R)^{-3} \approx 3/4.
  \end{equation}
  However, for planets which orbit the secondary star, the completeness is
  \begin{equation}
    f_{d2,c} = (1+\gamma_R^{-1})^{-3} \approx 3/4000.
  \end{equation}

@@

The model of Sec.~\ref{sec:model_1} is idealized, but it helps
develop intuition.
Now, we let the light ratio $\gamma_R = L_2/L_1$ vary across the population 
of star systems.
It does so because the underlying mass ratio $q=M_2/M_1$ varies.
We keep the primary mass fixed as $M_1$, which is also the mass of all single 
stars.

One parametrization for the distribution of 
binary mass ratios in a volume-limited sample is $f(q)\propto q^\gamma$.
[Duchene and Kraus 2013], fitting all the multiple systems of [Raghavan et al 
(2010)]'s Fig 16, quote $\gamma = 0.28\pm0.05$ for $0.7<M_\star/M_\odot<1.3$.
Examining only the binary systems of Rhagavan et al 2010, Fig 16, the 
distribution is roughly uniform,
\begin{equation}
{\rm prob}(q) =
\Bigg\{\begin{array}{lr}
c_q & 0.1 < q \leq 1  \\
0 & {\rm otherwise},
\end{array}
\label{eq:mass_ratio}
\end{equation}
provided we ignore the peak at high mass ratios for analytic convenience. $c_q 
= 1/9$ is the normalization.
We choose this latter mass ratio distribution.

For the mass-luminosity relation, for analytic convenience we assume 
$L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
As a simple check, we fit lines to mass-luminosity data\footnote{
    Data collected by Torres et al. [2009] for dwarfs earlier than 
    spectral class M, and Benedict et al. [2016] for dwarfs later than 
    spectral class M.
    To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
    luminosities, we interpolated over E. Mamajek's 
    table
	\texttt{pas.rochester.edu/\textasciitilde 
		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
	downloaded 2017.08.02}.
In log-log space, we let the intersection point of the lines float. The 
resulting fit parameters are available in a footnote\footnote{
	m lo: 1.8818719873988132~--~
	c lo: -0.9799647314108376~--~
	m hi: 5.1540712426599882~--~
	c hi: 0.0127626185389781~--~
	M at merge: 0.4972991257826812~--~
	L at merge: 0.0281260412126928.
}. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
$L=M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.

The final addition to our model is that we must assume a stellar mass-radius 
relation.
This affects the transit probabilities and signal sizes about the 
secondaries of binary star systems.
Since again, our main purpose is to develop intuition about observational 
biases, we approximate this as 
\begin{equation}
R\propto M.
\end{equation}
\begin{comment}
[Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
binary data,
\begin{equation}
R = 1.06 M^{0.945}
\label{eq:mass_radius}
\end{equation}
applies for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 
1991's stated bounds), where each quantity is given in solar units. 
\end{comment}

@@

Counting the number of desired stars, $N_s$ is the same as in 
Sec.~\ref{sec:model_1}. However
\begin{equation}
N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3
\end{equation}
for
\begin{equation}
d_{\rm max,d} =
d_{\rm max,s}\times (1+\gamma_R)^{1/2},
\end{equation}
where $d_{\rm max,s}$ is given by Eq.~\ref{eq:d_max}.

For a given binary system, $q$, $\gamma_R$, and thus $d_{\rm 
	max,d}(\gamma_R)$ are all random variables.
Since $N_d$ is a function of $d_{\rm max,d}$, 
the number of double star systems becomes a random variable, drawn from the 
ensemble of all possible transit surveys.
Applying the chain rule for probability density functions, the distribution 
function for the number of binary systems,
${\rm prob}(N_d)$, can be written 
\begin{equation}
{\rm prob}(N_d) = {\rm prob}(q(\gamma_R)) 
\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
\label{eq:prob_Nd_raw}
\end{equation}

@@

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/joint_prob_r_q.pdf}
	\end{center}
	\caption{The joint probability distribution of a binary star's position and 
		mass ratio in a magnitude-limited sample. This plot assumes $L\propto 
		M^{3.5}$. Note that the marginalized 
		mass ratio distribution is uniform at any given distance.}
	\label{fig:joint_prob_r_q}
\end{figure}

@@

The distribution function for the mass ratio of binaries in a magnitude 
limited sample can be found by marginalizing over the joint 
distribution for a binary star's position $r$ and mass ratio $q$.
Explicitly, using `ml' as a subscript for `magnitude-limited',
\begin{equation}
{\rm prob}_{\rm ml}(r,q) = {\rm prob}(q|r){\rm prob}_{\rm ml}(r),
\end{equation} 
where ${\rm prob}(q|r)$ is given by Eq.~\ref{eq:mass_ratio}.
One can show
\begin{equation}
{\rm prob}_{\rm ml}(r,q) \propto r^2 \quad{\rm if}\ r<d_{\rm max,d}(q)\ {\rm 
and}\ 0.1<q<1, 
\end{equation}
and otherwise zero. The resulting joint probability distribution is plotted in 
Fig.~\ref{fig:joint_prob_r_q}.
Marginalizing the joint distribution over distance, and assuming $L\propto 
M^\alpha$, one finds
\begin{align}
{\rm prob}_{\rm ml}(q) \propto (1+q^\alpha)^{3/2}\quad {\rm if\ 
}0.1<q\leq 1,
\label{eq:pdf_q_mag_limited}
\end{align}
and otherwise 0.
This marginalized distribution~--~the distribution function of mass ratios for 
binaries in a magnitude-limited sample~--~is plotted in 
Fig.~\ref{fig:q_distribn_mag_limited}.
Reworded, the probability of drawing a binary of mass ratio 
$q$ in a magnitude-limited sample scales as $d_{\rm max,d}^3(q)$.

Applying Eq.~\ref{eq:prob_Nd_raw} with the correct magnitude-limited 
distributions gives
\begin{equation}
{\rm prob}(N_d) \propto N_d^{2/3} 
\left( \left(\frac{N_d n_s}{N_s n_d}\right)^{2/3} - 1
\right)^{\frac{1-\alpha}{\alpha}}
\end{equation}
\begin{comment}
\begin{align}
{\rm prob}(N_d) \propto N_d^{2/3} &\left(\frac{3}{4\pi 
	n_d}\right)^{5/3} d_{\rm max,s}^{-5} \ \nonumber\\
&\times \left[ 
	\left(\frac{3 N_d}{4\pi n_d}\right)^{2/3}  d_{\rm max,s}^{-2} - 1
 \right]^{\frac{1-\alpha}{\alpha} }
\end{align}
\end{comment}
over the interval
\begin{equation}
N_d^{\rm lower} = \frac{4 \pi n_d}{3} (\sqrt{(0.1)^\alpha + 1} d_{\rm 
max,s})^3,
\end{equation}
\begin{equation}
N_d^{\rm upper} = \frac{4 \pi n_d}{3} (\sqrt{2}d_{\rm max,s})^3,
\end{equation}
and outside the stated interval ${\rm prob}(N_d) = 0$.

@@

The reason we derived all this is that we wanted to know the number of 
searchable binary stars.
For any given realization of the survey, this number is a random draw.
Computing the \textit{expected value} for the number of double star systems in 
the sample,
\begin{equation}
\langle N_d \rangle = \int_{0}^{\infty} N_d\, {\rm prob}(N_d) \,{\rm d}N_d.
\end{equation}
Letting $\alpha=3.5$, this can be expressed as
\begin{equation}
\langle N_d \rangle \propto \mathcal{R}^{-3/2}
\int_{0.1^{7/2} + 1}^{2} u^3 (u-1)^{-5/7} \,{\rm d}u ,
\label{eq:integral_Nd}
\end{equation}
where non-dimensional constants are not shown,
$u = \mathcal{R} N_d^{2/3}$,
and $\mathcal{R}$ is defined as
\begin{equation}
\mathcal{R} \equiv \left(\frac{3}{4\pi n_d}\right)^{2/3} d_{\rm max,s}^{-2}
= \left(\frac{n_s}{N_s n_d}\right)^{2/3}.
\end{equation}
The integral in Eq.~\ref{eq:integral_Nd} has a closed-form solution in terms 
of polynomials. Once evaluated, and after 
normalizing ${\rm prob}(N_d)$, to 4 significant digits it gives
\begin{equation}
\langle N_d \rangle = 1.590 \left(\frac{4\pi n_d}{3}\right) d_{\rm max,s}^3.
\end{equation}
This gives the useful expression
\begin{align}
\frac{\langle N_d \rangle}{N_s} &= 1.590\, \frac{n_d}{n_s}
\end{align}
Taking the sun-like star binary fraction of ${\rm BF} = 0.44$, this gives
$\langle N_d \rangle / N_s = 1.25$.
Compared to the binary-twin thought experiment of Sec.~\ref{sec:model_1}, in 
which the SNR-limited sample had $\approx 2.8$ times more binary systems than 
single systems, there are now $\approx 1.25$ times as many.
This is sensible, because the mean light ratio has decreased; it was 
previously 1, it is now $\langle \gamma_R \rangle \approx 0.27$.

@@

Returning to the original question, the total number of stars for any given 
realization of the survey is
\begin{equation}
N_{\rm stars,tot} = N_s + 2 N_d.
\end{equation}
Since $N_d$ is a random number, so is the total number of stars.
By assumption all primaries are of the desired type, but for secondaries only 
a fraction $f_{\rm desired}$ of them are of the desired type:
\begin{equation}
N_{\rm stars} = N_s + (1+f_{\rm desired}) N_d.
\end{equation}
Explicitly,
\begin{equation}
f_{\rm desired} = \frac{\int_{M_{\rm min}}^{M_{\rm max}} {\rm prob}_{\rm 
ml}(q)\,{\rm d}q }
{\int_{0.1M_\odot}^{M_\odot} {\rm prob}_{\rm 
        ml}(q)\,{\rm d}q}.
\end{equation}
For instance, if we define the ``desired type'' to be 
$0.7<M_\star/M_\odot<1.3$, we find $f_{\rm desired}\approx0.475$.

For a given survey draw, the number of planets of the desired type is 
then
\begin{equation}
N_{\rm planets} = 
\Gamma_{t,s}N_s + \Gamma_{t,d} N_d + 
f_{\rm des}\Gamma_{t,d}w_{d2} N_d,
\end{equation}
and the drawn ``true occurrence rate'' is
\begin{align}
\Gamma_t &= \frac{ N_{\rm planets}  }{ N_{\rm stars} } \\
\Gamma_t &= 
\frac{
	\Gamma_{t,s}N_s + (1+f_{\rm desired}w_{d2})\Gamma_{t,d} N_d
	}{
	N_s + (1+f_{\rm desired}) N_d
	}.
\end{align}

@@

The observer who has never heard of binary star systems wants to find the 
occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars 
with mass $0.7<M_\star/M_\odot < 1.3$ (with corresponding stellar radii and 
luminosities).
They detected $N_{\rm det}$ transit signals that appear to be from planets of 
this desired type.
The occurrence rate they then find, {\it i.e.} the apparent number of planets 
divided by the apparent number of ``stars'' (really, stellar systems) is
\begin{comment}
\begin{equation}
\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
\end{equation}
\end{comment}
\begin{equation}
\Gamma_{\rm a } =
\frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_{s,g}}.
\end{equation}

@@

The drawn number of planet detections of the desired type is 
\begin{equation}
N_{\rm det} = 
N_{{\rm det},s} 
+ 
N_{{\rm det},d}.
\end{equation}
As in Sec.~\ref{sec:model_1},
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
\label{eq:N_det_s_model2}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, $f_{s,g}$ is the geometric transit probability, 
and $f_{s,c}$ is the fraction of these transiting planets that are observed 
with signal to noise greater than the minimum detection threshold (the 
completeness). 
For our SNR-limited, varying-$\gamma_R$ survey, $f_{s,g} = R_1/a\equiv f_g$, 
and $f_{s,c}=1$.

For planets detected in binary systems, only a fraction will be ``of the 
desired type'', since only a fraction of the stars are of the desired type.
This is the same factor $f_{\rm desired}$ as previously discussed. Accounting 
for this and the independent geometric and completeness probabilities,
\begin{equation}
N_{{\rm det},d}
= 
N_d \Gamma_{t,d} (f_{d1,g}f_{d1,c} + f_{\rm desired}w_{d2}f_{d2,g}f_{d2,c}).
\label{eq:N_det_d_model2}
\end{equation}

For any given system,
\begin{align}
f_{d1,c} &= (1+\gamma_R)^{-3} \\
f_{d1,g} &= f_{s,g} \\
f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \\
f_{d2,g} &= f_{s,g}q^{2/3}
\end{align}

\subsection{Errors on derived occurrence rates}

Recall $X_\Gamma = \Gamma_t / \Gamma_a$. Let $f_{\rm desired} = f_{\rm d}$ for 
brevity.
Combining the above results, for any draw from the ensemble of transit 
surveys the necessary correction to the derived occurrence rate is
\begin{align}
X_\Gamma =
&\frac{\Gamma_{t,s}N_s + (1+f_{\rm d}w_{d2})\Gamma_{t,d}N_d }{N_s + (1+f_{\rm 
d})N_d}
\nonumber \\
&\times
\frac{(N_s + N_d)f_{s,g}}{N_s \Gamma_{t,s} f_{s,g} + N_d \Gamma_{t,d}
      (f_{d1,g}f_{d1,c} + f_{\rm d}w_{d2}f_{d2,g}f_{d2,c})  }.
\end{align}
Letting $\beta = N_d/N_s$, 
\begin{align}
X_\Gamma =
&\frac{1+\beta}{1 + \beta(1+f_{\rm d})} 
\nonumber \\
&\times
\frac{ 1+ \beta (\Gamma_{t,d}/\Gamma_{t,s}) (1+f_{\rm d}w_{d2}) }
{1 +\beta(\Gamma_{t,d}/\Gamma_{t,s})(f_{d1,c} + f_{\rm 
d}w_{d2}f_{d2,g}f_{d2,c}/f_{s,g})}.
\end{align}

One typically assumes that $\Gamma_{t,s}=\Gamma_{t,d}$, and that the 
fraction of secondaries of the desired type with planets, $w_{d2}$ can be set 
to 1.
If so,
\begin{align}
X_\Gamma =
&\frac{1+\beta}
{1 +\beta(f_{d1,c} + f_{\rm d}f_{d2,g}f_{d2,c}/f_{s,g})}.
\label{eq:model2_xgamma_simple}
\end{align}

If $\gamma_R \ll 1$, $f_{d1,c}\approx 1$, $f_{d2,c}\ll f_{d1,c}$, and the 
number of detections from secondaries of binary systems is negligible.
In this limit, binarity does not affect the derived occurrence rate:
\begin{equation}
X_\Gamma = 1 \quad ({\rm limit\ of}\ \gamma_R \ll 1).
\end{equation}

To assess whether it is reasonable to take this limit, we want the mean
value of $X_\Gamma$, or the median, or else some other summary statistic.
Recall that all the terms in Eq.~\ref{eq:model2_xgamma_simple} are
random variables, and note that they are inter-dependent.
This latter fact impedes analytic progress.

Running a Monte Carlo simulation in which we fix the planet and primary star 
properties, draw the binary mass ratio from Eq.~\ref{eq:pdf_q_mag_limited}, 
and take 
the luminosity as $L\propto M^{3.5}$, we find 
a mean of $\langle X_\Gamma \rangle ={\bf VALUE HERE}$,
and a median of ${\rm median}(X_\Gamma) ={\bf VALUE HERE}$.
The distribution is shown in Figure BLAH.

\begin{comment}
\subsection{How many planets are in the sample?}

@@

\end{comment}
